### 2021-01-28 input/proposal.tex
wip

A reactive logical partitioning the cost of which actually depends on
its usage.

@@

@@

>>>>

In this section, we introduce \NAME, (\TODO{stands for}) a reactive
protocol for logical partitioning the cost of which actually depends
on its usage: when the system becomes quiescent, processes eventually
converge to their respective partition and do not require further
communication afterwards.

@@

This section describes \NAME's components by starting from the
simplest case of add-only partition, to the most complex case of
dynamic partitions in dynamic networks.

\TODO{Probably better to pair operation as ``ins, del'' and ``add, rem''?}

\paragraph{Static network with add-only.}
At any time, a process can become a source, hence \TODO{creating} a
new partition in the system. This partition eventually includes all
processes that are closer from this new source than any other else. We
described such protocol in Section~\ref{sec:background}. Processes
naturally converge towards their respective best partition by only
piggybacking a monotonically increasing distance in forwarded
messages.  \TODO{Traffic of each partition is contained to the
  partition.}

\paragraph{Static network with dynamic partitions.}
At any time, a source can revoke its self-appointed status of source,
hence \TODO{removing} its partition of the system. All processes that
belong to this partition must choose another partition to belong to.
To handle dynamic partitions, processes require two additional
components.

\noindent \textbf{First,} since the number of partitions does not
monotonically increase in the system any longer, they require a vector
of versions that monotonically increases over \TODO{add and del}
operations. This allows processes to quickly discard stale messages
saving bandwidth. For instance, a process that received a message
originated at Process $a$ with a version $2$ knows that any message
originated at Process $a$ with a version $1$ are stale system-wide; it
must not forward it. As consequence, vectors of versions also ensure
termination, for messages \TODO{add and del} cannot follow each other
in an infinite loop. In terms of traffic, this only requires each
\TODO{delete} message to piggyback a source identifier and version
number.

\noindent \textbf{Second,} processes need to maintain a local view of
their neighbors' partition.  Figure~\ref{fig:del} shows that
\TODO{deletes} may not propagate properly to all processes in
need. The path leading to processes delivering $add$ may be broken due
to concurrent deliveries.  To solve this issue, each process is able
to know with certainty that each of its neighbor does not need a
delete (either because it is in another partition, or because another
path exists from the source to this neighbor). \TODO{More, figures
  etc.} In terms of traffic, this requires to acknowledge each message
in a way or another (acknowledgement messages or included in each
add-del message) using FIFO communication channels. The propagation of
delete messages is limited to the partition it targets plus uncertain
processes.

\noindent \textbf{Finally,} the delete must trigger another
competition amongst neighboring partitions. These \TODO{add}
messages operate normally and fill gaps left open by deletes.

\paragraph{Networks with new communication links.}
Adding new communication links to the network may create shortcuts
between processes. Both processes must send their current best
partition to each other. Upon receipt, they act normally: if a process
finds out that the received partition is closer than its current one,
it delivers it which in turns also triggers another competition
amongst neighbors due to forwarding.

Joining the network is equivalent to add as many communication links
as necessary between the joining process and its new neighbors.

\paragraph{Dynamic networks with dynamic partitions.}
When removing a communication link between two processes does not
break any active path, because neither distances of processes depend
on the other, then nothing needs to be done. \NAME have no overhead.
Unfortunately, when a process' distance depends on the other process,
the protocol becomes much more complex. Indeed, this requires to
\TODO{undo} all add messages originated from this process. \TODO{Why
  distance is not sufficient ?} \TODO{Why not use this from the first
  step then, or why other steps are still relevant?}

\paragraph{A last optimisation}

\TODO{Carry information that is useless. such as order of processes in
  paths, and identities of processes themselves.}

@@

\TODO{To rework, for there are more components now.}
